Formalization and Automation of Variants of G\"{o}del's Ontological Proof of God's Existence

### 2013-08-16 Abstract/GodProof-Abstract.tex
Improve abstract

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
%
In our recent work we did employ the following axioms\footnote{Say where they come from and how they relate to G\"odel.}: (1) Any property strictly implied by a positive property is positive,
(2) A property is positive if and only if its negation is not positive, and (3) the property of being God-like is positive. Moreover, the following definitions are employed:
(A) x is God-like if and only if x incorporates all positive properties, (B) a property P is essential for x (and essence of x) if and only if P is a property of x and every property Q that x has is strictly implied by P, (C) x necessarily exists if and only if every essence of x is necessarily exemplified.

From these axioms and definitions we then infer:
\begin{itemize}
\item Positive properties are eventually exemplified.
\item Eventually God exists.
\item If x is God-like, then the property of being 
God-like is an essence of x.
\item Necessarily God exists.
\end{itemize}

@@

In the literature \cite{ToDo} slightly different versions of the axioms and definitions exist 
and derivations of the above mentioned claims are presented with various degrees
of detail and formality. The above variant of G\"{o}del's proof has now been 
constructed for the first-time
with the utmost degree of detail and formality; cf.~\cite{ToDo:GitHubRepository}. The following has been done:
\begin{itemize}
\item A detailed natural deduction proof.
%
\item A formalization of the axioms and theorems in the TPTP THF format \cite{J22}.
%
\item Automatic verification of the consistency of the axioms and 
definitions with Nitpick \cite{Nitpick}.
%
\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.

@@

G\"{o}del's proof is challenging to formalize and verify 
because it requires very expressive logical languages with 
modal operators (\emph{possible} and \emph{necessary}) and higher-order quantifiers. 
Our computer-assisted formalizations rely on an embedding of the modal logic S5 
into classical higher-order logic with Henkin 
semantics \cite{J23} and 
employed recently developed interactive and automated deduction tools designed for this logic.

>>>>

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof. Our formalization employs the following axioms\footnote{Say where they come from and how they relate to G\"odel.}: (1) Any property necessarily implied by a positive property is positive;
(2) A property is positive if and only if its negation is not positive; (3) the property of being God-like is positive; (4) positive properties are necessarily positive; and (5) necessary existence is a positive property.
Moreover, the following definitions are employed:
(D1) x is God-like if and only if x has every positive property, (D2) a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, (D3) x necessarily exists if and only if every essence of x is necessarily exemplified.

From these axioms and definitions we then infer:
\begin{enumerate}
\item Positive properties are possibly exemplified.
\item Possibly God exists.
\item If x is God-like, then the property of being 
God-like is an essence of x.
\item Necessarily God exists.
\end{enumerate}

@@

The above variant of G\"{o}del's proof has now been 
constructed for the first-time
with the utmost degree of detail and formality; cf.~\cite{ToDo:GitHubRepository}. The following has been done:
\begin{itemize}
\item A detailed natural deduction proof.
%
\item A formalization of the axioms and theorems in the TPTP THF format \cite{J22}.
%
\item Automatic verification of the consistency of the axioms and 
definitions with Nitpick \cite{Nitpick}.
%
\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.

@@

G\"{o}del's proof is challenging to formalize and verify 
because it requires very expressive logical languages with 
modal operators (\emph{possibily} and \emph{necessarily}) and higher-order quantifiers. 
Our computer-assisted formalizations rely on an embedding of the modal logic S5 
into classical higher-order logic with Henkin 
semantics \cite{J23} and 
employed recently developed interactive and automated deduction tools designed for this logic.
